Speed and strength of epidemic intervention

### 2019-07-25 doc/speedstrength.tex
more work

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}

@@

\section{Examples}

@@

As our baseline scenario (\figref{example}A), we assume that early transmission is responsible for 23\% of the total transmission \citep{hayes2006amplified}.
Based on the growth rate estimate ($0.452 \textrm{year}^{-1}$; \figref{example}B) from a HIV prevalence data in pregnant women in South Africa \citep{barron2013eliminating}, our baseline reproduction number is $3.17$.
Further increasing the amount of early transmission reduces the mean generation time, which in turn decreases the reproduction number (\figref{example}C).

\begin{figure}
\includegraphics[width=\textwidth]{../figure/figure1.pdf}
\caption{This is not a figure.}
\figlab{example}
\end{figure}

\subsection{Univarsal testing and treatment}

\subsection{Test and treatment - HIV}

\begin{itemize}
	\item Can active testing and treatment stop the epidemic?
	\item Explain different scenarios? assumptions?
	\item Early vs late transmission (change proportion of early transmission)
	\item Compare required strength and speed of intervention required to control the disease
	\item Early transmission matters for condom
	\item Early transmission doesn't matter so much for target and treat
	\item Think about speed vs strength axis
\end{itemize}

\subsection{Asymptomatic infection}

\begin{itemize}
	\item See ``Factors that make an infectious disease outbreak controllable'' by Fraser et al.
	\item Our framework generalizes their work
	\item Timing of target intervention (Fraser works on this stuff and did a clinical trial on this; unpublished)
\end{itemize}

\subsection{Maybe one more example}

@@

\begin{itemize}
	\item Talk about generation time and reinterpret Eaton and Hallet?
	\item Short generation time -> teating and treating will help less but lower epidemic strength
	\item High generation time -> high proportion prevented but higher R.
\end{itemize}

\begin{itemize}
	\item r and R have more in common than we think
	\item Why test and treat predictions are robust to assumptions about transmission
\end{itemize}

>>>>

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
\usepackage[top=1in,left=1in, right = 1in, footskip=1in]{geometry}

@@

\section{Intervention strategies for HIV}

@@

\begin{figure}[!t]
\includegraphics[width=\textwidth]{../figure/figure1.pdf}
\caption{This is not a figure.}
\figlab{example}
\end{figure}

As our baseline scenario (\figref{example}A), we assume that early transmission is responsible for 23\% of the total transmission \citep{hayes2006amplified}.
Based on the growth rate estimate ($0.452 \textrm{year}^{-1}$; \figref{example}B) from a HIV prevalence data in pregnant women in South Africa \citep{barron2013eliminating}, our baseline reproduction number is $3.17$.
Further increasing the amount of early transmission reduces the mean generation time, which in turn decreases the reproduction number (\figref{example}C);
therefore, we expect stronger (but not necessarily faster) intervention to be required in order to control the disease when there is more early transmission.

\swp{Probably don't need a figure for this?}
First, consider a condom intervention, which is approximately 80\% effective in reducing HIV seroconversion \citep{weller2002condom}.
Assuming that ?\% of the sexually active population uses condom consistently,
we can model this intervention as a constant function: $\tsub{L}{condom} = ???$.
In this case, the strength of condom intervention is always constant regardless of the amount of early proportion.
Therefore, we expect condom intervention to be able to control the spread of the HIV when more than ???\% of transmission is caused by early transmission (i.e., when the intervention strength is greater than the epidemic strength).

\begin{figure}[!t]
\includegraphics[width=\textwidth]{../figure/figure2.pdf}
\caption{This is not a figure.}
\figlab{test}
\end{figure}

On the other hand, consider a test-and-treat strategy in which infected individuals are identified and put to clinical care (e.g., antiretroviral therapy).
We can model this intervention $\tsub{L}{test}$ as a saturating function that begins at 1 and plateaus at a sufficiently high value (such that identified individuals are effectively no longer infectious);
in particular, we assume that all individuals will be identified and put to treatment within the first 5 years since their infection (\figref{test}A).
Under this assumption, the strength of intervention decreases as the proportion of early transmission increases but is always higher than the epidemic strength; 
however, it is unclear why this should be the case.

Using the speed-based decomposition allows us to understand the problem much more clearly.
The hazard rate $\tsub{h}{test}$ of this intervention starts at 0 (because infected individuals don't have a way to know that they have HIV when they're infected) but peaks very fast (because individuals are actively getting tested and treated); 
the hazard rate goes down soon after the peak and reaches 0 after about 5 years because infected individuals who have access to medical care would have been identified within the first 5 years or so.
The speed of epidemic stays constant regardless of the proportion of the early transmission because epidemic speed corrresponds to the observed growth rate.
The speed of intervention stays roughly constant but slightly increases as proportion of early transmission increases \swp{because?}.
This framework clearly demonstrates that test-and-treat intervention will be effective regardless of the proportion of early transmission.

@@
